Linting Linearity in Core/System FC

### 2023-01-16 report.tex
Expand Haskell section introduction

\begin{itemize}
    \item Ideia dos linear types
    \item Calculo lambda linear (como forma de falar da linearidade)
    \item Gestão de recursos
    \item Rust (core is linear lambda calculus?)
\end{itemize}

@@

Haskell is a functional programming language defined by the "Haskell
Report"\cite{jones1999haskell,marlow2010haskell} and whose \emph{de-facto}
implementation is GHC, the Glasgow Haskell Compiler\cite{GHC}. Haskell is a
\emph{lazy}, \emph{purely functional} language, i.e., functions cannot
have side effects or mutate data, and contrary to many programming languages,
arguments are \emph{not} evaluated when passed to functions, but rather are only
evaluated when the value is needed.

>>>>

% \begin{itemize}
%     \item Ideia dos linear types
%     \item Calculo lambda linear (como forma de falar da linearidade)
%     \item Gestão de recursos
%     \item Rust (core is linear lambda calculus?)
% \end{itemize}

@@

Haskell is a functional programming language defined by the "Haskell
Report"\cite{jones1999haskell,marlow2010haskell} and whose \emph{de-facto}
implementation is GHC, the Glasgow Haskell Compiler\cite{GHC}. Haskell is a
\emph{lazy}, \emph{purely functional} language, i.e., functions cannot
have side effects or mutate data, and contrary to many programming languages,
arguments are \emph{not} evaluated when passed to functions, but rather are only
evaluated when the value is needed. The combination of purety and laziness is
unique to Haskell among mainstream programming languages...

Haskell is a large language but its small core is based on a typed lambda
calculus. As such, there exist no statements and computation is done simply
through the evaluation of functions. Besides functions, one can define types and
their constructors and pattern match on said constructors. Function application
is denoted by an empty space (\textbf{f~a} means \textbf{f} applied to
\textbf{a}) and pattern matching is done with the \textbf{case} keyword followed
by the enumerated alternatives. All variable names start with lower case and
types start with upper case (excluding type variables). To make explicit the
type of an expression, the \textbf{::} operator is used (e.g.
$\textbf{f~::~Int}\to \textbf{Bool}$ is read \textbf{f} \emph{has type} function
from \textbf{Int} to \textbf{Bool}).

Because Haskell is a pure programming language, input/ouptut side-effects are
"made pure" by modelling them on the type level through the
higher-kinded\footnote{\textbf{IO} has kind $\textbf{Type}\to\textbf{Type}$,
that is, it is only a type after another type is passed as a parameter (e.g.
\textbf{IO~Int}, \textbf{IO~Bool}); \textbf{IO} by itself is a \emph{type
constructor}} type \textbf{IO}. Some of the example programs will look though as
if they had statements, but, in reality, the sequential appearence is just
syntatic sugar to an expression using Monadic operators. The main take away is
that computations that do I/O may be sequenced together with other operations
that do I/O while retaining the lack of statements and the language purity
guarantees.

% TODO: Perhaps elaborate a bit more on the constructors and pattern matching?

As an example, consider these functions that do I/O and their types. The first
opens a file by path and returns its handle, the second gets the size of a file
from its handle, and the third closes the handle. It is important that the
handle be closed exactly once, but currently nothing enforces that usage policy.

\begin{code}
openFile :: FilePath -> IOMode -> IO Handle
hFileSize :: Handle -> IO Integer
hClose   :: Handle -> IO ()
\end{code}

The following function makes use of the above definitions to return the size of
a file given its path.
\begin{code}
countWords :: FilePath -> IO Integer
countWords path = do
    handle <- openFile path ReadMode
    size   <- hFileSize handle
    return size
\end{code}
% hClose handle
